Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2022-06-24 elsarticle/elsarticle.tex
updates
add line re. wang
Add animal data

%% or include affiliations in footnotes:
\author[mymainaddress]{Matz Andreas Haugen\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
% \ead[url]{www.matzhaugen.com}
\ead{matzhaugen@gmail.com}

\author[mymainaddress]{Dorothea Gilbert}

\address[mymainaddress]{Independent Researcher, Oslo, Norway}
% \address[mysecondaryaddress]{Independent Researcher, Oslo, Norway}

@@

 At this limit the relative growth rate approaches infinity as $Y \rightarrow 0$, which seems odd under the hypothesis that the pathogen has just started spreading. 
 The Gompertzian limit also implies a decreasing relative growth rate from the very first time point, which under the SIR model seems unlikely given the large pool of susceptible people in the beginning. 
 One would rather expect a near-constant relative growth rate in the beginning due to a disease propagation stage. 
 Rypdal and Rypdal \cite{Rypdal2020} suggest that the decreasing relative growth rate is caused by social and political mitigating efforts, but these hardly justify such coherent and consistent mortality characteristics across countries.

@@

\section{An alternative theory for observed mortality}
\label{seq:alt}
An alternative line of reasoning that does not rely on the framework of communicable diseases is that the biosphere was perturbed by an external stressor, initiating a stress response to eventually bring mortality rates back to stability. 
This could explain the immediate dampening of mortality growth observed. 
It could also explain the lack of correlation between population density and mortality (or infection) rates observed by some \citep{Hamidi2020,Hamidi2020a,Carozzi2020,Arpino2020,khavarian2021high,barak2021urban}.

>>>>

%% or include affiliations in footnotes:
\author[mymainaddress, mysecondaryaddress]{Matz Andreas Haugen\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
% \ead[url]{www.matzhaugen.com}
\ead{matzhaugen@gmail.com}

\author[mysecondaryaddress]{Dorothea Gilbert}

\address[mymainaddress]{Enerhauggata 1, 0651, Oslo, Norway}
\address[mysecondaryaddress]{Independent Researcher, Oslo, Norway}

@@

 At this limit the relative growth rate approaches infinity as $Y \rightarrow 0$, which seems odd under the hypothesis that the pathogen has just started spreading. 
 The Gompertzian limit also implies a decreasing relative growth rate from the very first time point, which under the SIR model seems unlikely given the large pool of susceptible people in the beginning. 
 One would rather expect a near-constant relative growth rate in the beginning due to a disease propagation stage. 
 Rypdal and Rypdal \cite{Rypdal2020} suggest that the decreasing relative growth rate is caused by social and political mitigating efforts, but these hardly justify such coherent and consistent mortality characteristics across countries. In fact, Wang et al. \cite{wang2012richards} showed that values of $\theta<1$ are not biologically realistic under a similar SIR model as the one given above.

@@

\section{An alternative theory for observed mortality}
\label{seq:alt}
An alternative line of reasoning that does not rely on the framework of communicable diseases is that the biosphere was perturbed by an external stressor, initiating a stress response to eventually bring mortality rates back to stability. 
This could explain the immediate dampening of mortality growth observed not explainable by a SIR model. 
It could also explain the lack of correlation between population density and mortality (or infection) rates observed by some \citep{Hamidi2020,Hamidi2020a,Carozzi2020,Arpino2020,khavarian2021high,barak2021urban}, in addition to the widespread morbidity and mortality spikes seen during spring 2020 among birds \citep{Fischer2021,NINA2020}, dogs \citep{PeriseBarrios2021}, rabbits \citep{Duff2020,Hu2020,Fukui2021}, elephants \citep{Aarde2021}, and horses \citep{Kambayashi2021,King2020,Lu2020,CastilloOlivares2020}.
